---

output:
  word_document:
    reference_docx: word_styles.docx
editor_options: 
  chunk_output_type: inline
  
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())

```
Crime in schools
Crime in Mexico schools
SNA and Crime - previous methods and descriptions
My methods
Results
conclusion

```{r echo=FALSE, message=FALSE, warning=FALSE}
library("statnet")
library("dplyr")
library("knitr")
library("tidyr")
library("ggplot2")
library("readxl")
library("reshape2")
library("stringr")
library("RColorBrewer")
library("flextable")
library("tibble")
library("tidyverse")
library("network")
library("statnet")
library("UserNetR")
library("sna")
library("corrplot")
library("devtools")



```

## create network, nodes and edges

Creating a network for analysis actualizes connections between nodes and edges. These entities are used to create network objects that represent adjacency matrices, or sociomatricies, which are square matricies with nodes as the network's rows and columns and the binary connections (0, 1) as their edges. At its simplist connection, for example, a link between two nodes is represented by a 1, 0 otherwise. Sometimes networks can be directed where a node sends information to a recieving node, but the information is not reciprocated. In our case, our data is undirected and represents two groups, crimes and schools. ADD DESCRIPTION OF SOCIOMATRIX?? Density of the network = 0.03590093 

```{r echo=FALSE, message=FALSE, warning=FALSE}

crime150 <- read_xlsx("C:/Users/LW/Box/Mexico City Police 2020/data/buffer_150_crime3.xlsx")

#crime columns
delito_espesifico <- crime150$delito
delito_vec <- as.array(crime150$categoria_)
delito <- crime150$categoria_
crime_spot <- crime150$calle_hech
crime_neighbrhd <- crime150$colonia_he
crime_region <- crime150$alcaldia_h

length(unique(delito_espesifico))

#school columns
escuela_todo <- crime150$escuela_nombre
nivel <- (crime150$escuela_nivel)
nombre <- crime150$escuela
school_address <- crime150$domicilio

# length(unique(nivel))
# table(delito_espesifico)
# length(unique(delito_espesifico))

# mtx1 <- matrix(crime150, nrow = length(unique(delito_espesifico)), ncol = length(unique(nivel)))

mtx_crime150 <- data.matrix(crime150)
# mtx_crime150
# glimpse(crime150)
# crime <- model.matrix(~ delito_espesifico + nivel -1)

new_crime <- crime150[, c("categoria_","delito", "escuela_nivel", "calle_hech", "colonia_he", "alcaldia_h", "escuela", "domicilio")]
crime_mtx <- data.matrix(new_crime)
# crime_mtx
summary(crime_mtx, print.adj = FALSE)

mtx <- crime150[, c("delito", "escuela_nivel")]
# glimpse(mtx)
mtx_fin <- mtx %>% drop_na() #[complete.cases(mtx), ] #drop na's #na.omit(mtx) 
# glimpse(mtx_fin)



bin_crime <- as.data.frame.matrix(table(mtx_fin))
mtx_full <- as.matrix(bin_crime)
bin_crime <- as.matrix((bin_crime>0) +0)

crime_net <- network(bin_crime)
# dim(mtx_fin)

# summary(crime_net)

my.colors <- c(rep("red"), rep("blue"))
gplot(crime_net, gmode="twomode", 
      label=c(rownames(delito_espesifico),colnames(nivel)),
      vertex.cex=2,
      usearrows=TRUE,
      label.col = my.colors,
      main = "Crime to School Network: Blue = School, Red = Crime",
)
```


## Centrality Measures

Looking at the centrality measures of a network helps to identify various ways in which a node is connected to other nodes. There are 3 basic types of centrality measures including degree centrality, closeness centrality, and betweenness centrality. Degree centrality measures the number of connections a node is to other nodes within the network (Wasserman, Faust, 1994). Closeness centrality meausures how close a node is to all the other nodes in the network based on their distance apart, or how long it takes to get from one node to another. Betweenness centrality measures how many nodes a particular node sits between, for example, how central is this node for facilitating information flow within a network, are they an essential stop or can they be skipped?

The degree centrality for the top three school nodes is, PRIMARIA GENERAL (d = 198) translates to "public elementary school", PREESCOLAR GENERAL(d = 160) translantes to "public preschool", and SECUNDARIA GENERAL (d = 116) translates to "public middle school". The degree centrality for the top crime nodes is VIOLENCIA FAMILIAR (d = 30) translates to "domestic violence", ROBO A TRANSEUNTE EN VIA PUBLICA CON VIOLENCIA (d = 28) translates to "mugging", ROBO A NEGOCIO SIN VIOLENCIA (d = 24) translates to "non-violent business robbery", and AMENAZAS (d = 24) translates to "threats." With this information we can assume that public schools have a higher number of crimes in their neighborhood, and the top crimes in a school's a 150m radius are domestic "violence", "mugging", "non-violent business robbery", and "threats." Overall, the top ten nodes with the highest degree measures were schools, which is inline with the type of data we are using as schools in the center of their own 150m radius. 

The closeness centrality measures for the top three school nodes are "public elementary school", (c = 0.64), "public preschool" (c = 0.55), and "public middle school"(c = 0.47). The top crime centrality measure are "domestic violence" (c = 0.50), "mugging" (c = 0.49), "business robbery" (c = .048), and "threats" (c = 0.48). The top school and crime centralities are the same, but their order of greater closeness starts with "public elementary school", "public preschool", "domestic violence", and then "mugging". We can assume that these top nodes happen in close proximity to each other.

The top betweenness centrality measures for school nodes are also "public elementary school" (b = 11055.1143),"public preschool" (b = 6530.8892), and "public middle school" (b = 3473.3928). The top betweenness scores for crime slightly changed and their measures are, "domestic violence" (b = 1406.7534), "mugging" (b = 977.0407), and ROBO DE VEHICULO DE SERVICIO PARTICULAR SIN VIOLENCIA (b = 889.8290) translates to "non-violent theft of a private vehicle". The top 5 betweeness nodes overall are schools, meaning that schools are a central componenet for crimes in the neighborhood. This is expected because of the nature of the data and how the radius around the schools was set-up to collect the crime data.

```{r echo=FALSE, message=FALSE, warning=FALSE}

require(sna)

# summary(crime_net)
# crime_net$val

print("Degree Centrality: Crimes")
d.cent <- degree(crime_net) #[1:126, 1:126])
# d.cent
# table(d.cent)
names(d.cent) <- crime_net %v% "vertex.names"
sort.d <- sort(d.cent)
sort.d[149:163]
top.d <- names(sort(d.cent))[157:163]
# Draw the graph with node sizes proportional to betweenness centralitiy scores:
gplot(crime_net,  label=top.d, vertex.col = my.colors, usearrows=TRUE, vertex.cex = .05 * d.cent) + #label=names(d.cent), #label=names(d.cent),
title(main = "Degree Centrality:Crime Network")



print("Closeness Centrality: Crimes")
c.cent <- closeness(crime_net, gmode = "digraph")
# c.cent
names(c.cent) <- crime_net %v% "vertex.names"
# sort(c.cent)
sort.c <- sort(c.cent)
sort.c[153:163]
top.c <- names(sort(c.cent))[155:163]
# note there are all 0s for closeness centrality
# Draw the graph with node sizes proportional to closeness centralitiy scores:
gplot(crime_net,  label=top.c, vertex.col = my.colors, usearrows=TRUE, vertex.cex = 5 * c.cent) + # label=names(c.cent),
title(main = "Closeness Centrality:Crime Network")

print("Betweeness Centrality: Crimes")
b.cent <- betweenness(crime_net, gmode = "digraph")
# table(b.cent)
names(b.cent) <- crime_net %v% "vertex.names"
sort.b <- sort(b.cent)
sort.b[153:163]
top.b <- names(sort(b.cent))[155:163]
# Draw the graph with node sizes proportional to closeness centralitiy scores:
gplot(crime_net, label=top.b, vertex.col = my.colors, usearrows=TRUE, vertex.cex = .05 * b.cent) + #  
title(main = "Betweenness Centrality:crime Network")

```

## Hierarchical Clustering and structural distances
```{r echo=FALSE, message=FALSE, warning=FALSE}
# In-class: CONCOR in R

# Okay, let's see how to do something similar using the programs in the
#  "sna" package.
# Why would you (sometimes) prefer my programs (above) to the sna package? The answer,
#  if I understand things right, is that my concorFirst program has more options
#  than does sna for computing correlations on stacks.  Even if that is true, however,
#  sna has some nice features. (By the way, the "stats" package is built into R, so you
#  do not need to install it-it's already there.) 

require(sna)
# Structural equivalence distance on the basis of (Euclidean) distance:
eucl_str <- sedist(bin_crime, diag=FALSE, method="euclidean")
round(eucl_str[1:10,1:10], digits = 3) #
# The above is identical to page 372 in Wasserman-Faust (euc. distances)!

corr_str <- sedist(bin_crime, diag=FALSE, method="correlation")
round(corr_str[1:5,1:5], digits =3)
# The above is identical to the correlations on page 373!

# Let's try some hierarchical clustering:
require(stats)
plot(hclust(as.dist(eucl_str), method = "complete"))
# The above is identical to Fig. 9.7 (page 383) in Wasserman-Faust.
# Looking at the plot (eyeballing it), if we wanted to divide this tree into (say)
#   four blocks, we could do it with a horizontal line at a height of 
#   about 4.4 (intercept = 4.4, slope = 0); let's make it a thick red line:
abline(4.4, 0, col="red", lwd = 2)
# ("4.4" refers to a distance with reference to the distance matrix in page372.)
# Now looking at the tree ("cluster dendrogram) in the plot, this gives us
#  the following blocks:
#
#   7, 21
#   1, 2, 8, 16, 6, 12, 14, 11, 17
#   10, 18                          # Just as shown on p. 382 of Was-Faust!
#   13, 19, 4, 20, 3, 15, 5, 9      #   (the "first" blocking on that page)


# Now let's do an MDS of our network data. We need to convert correlations to distances.
# Here is a seat-of-the-pants way to do this:
require(MASS)

rough.dists <- (1 - corr_str) / 2
rough.dists <- dist(unique(rough.dists)) 
# Thus, a correlation of -1 becomes a "rough.dist" of 1, and
#     a correlation of +1 becomes a "rough.dist" of 0.
result <- isoMDS(rough.dists)
coordinates <- result$points
plot(coordinates[,1], coordinates[,2], type="p", main="Just like p. 387 in Wasserman-Faust!")
text(coordinates[,1], coordinates[,2], labels = 1:37, pos = 2)


```

##Automorphic equivalence

```{r echo=FALSE message=FALSE, warning=FALSE}

# This script illustrates the "maxsim" algorithm for a data-dredging
#  approach to finding automorphic equivalence classes. The example
#  is Fig. 13 (p. 30) in S. Borgatti & M. Everett, "Notions of
#  Position in Social Network Analysis," Sociological Methodology 22
#  (1992): 1 - 35. url = https://www.jstor.org/stable/270991


# Compute shortest paths:
require(sna)
# gd <- geodist(bin_crime)
gd <- geodist(crime_net)
gdist <- gd$gdist
# Take a look at the shortest-path matrix:
dim(gdist)

# Take the inverse (to create distances to similarities),
#  but keep 0's on the diagonal:
inv.gdist <- 1 / gdist
diag(inv.gdist) <- 0
# Take a look at the inverse path distances:
round(inv.gdist, digits = 3)
#
# Write a program to compute the Euclidean distance between
#   two strings of numbers:
euc.distance <- function(x,y) {
  dist <- sum((x - y)^2)
  dist <- sqrt(dist)
  return(dist)
}


# Here is the "work" of the "maxsim" program: in comparing two
#  rows, we sort each row in the same order, WITHOUT regard to
#  matching individuals. Then we compute the distance between
#  their "profiles." We'll use AE for automorphic equivalence
#  distance:
AE <- matrix(0, nrow = 163, ncol = 163)
rownames(AE) <- colnames(AE) <- rownames(gdist)
for (i in 1:10) {
  for (j in (i + 1):163) {
    ordered.i <- sort(inv.gdist[i,])
    ordered.j <- sort(inv.gdist[j,])
    AE[j,i] <- AE[i,j] <- euc.distance(ordered.i, ordered.j)
  }}

# Same as in my handout. (We don't change anything if we multiply
#   all numbers by 10):
round(10 * AE, digits = 2)

# Hierarchical clustering using "complete link" clustering:
hc <- hclust(d = as.dist(AE), method = "complete")
plot(hc)
abline(4.4, 0, col="red", lwd = 2)

# This gives the automorphic-equivalence groups for
# Borgatti& Everett's Fig. 13, viz:
#  {Mary, Bill}
#  {Larry, Moe}
#  {Curly}
#  {Jane, Sally, Charles, Linda, John, Michael}
# As Borgatti & Everett note on their p. 31, the 6 actors in
#  the last-listed set are all automorphically equivalent
#  to one another because they are all peripheral.

# Finally, let's do a plot of the network that shows the
#  automorphic equivalence classes:
# memberships <- rep(0, 163)
# memberships[c(5,7)] <- 1
# memberships[c(3,9)] <- 2
# memberships[6] <- 3
# memberships[c(11,10,8,4,1,2)] <- 4
vertex.cols <- c("lightblue","green","brown","yellow")
label.cols <- c("black","white","white","black")
require(sna)
gplot(gdist, usearrows=FALSE, vertex.cex=1.8,
      label.cex=.8, vertex.col=vertex.cols[AE],
      label.col=label.cols[AE], label.pos=5, gmode="kamadakawai")
title(main="Automorphic Equivalence Classes", sub="Borgatti & Everett, Soc Meth 1992, p. 30")

```

## Duality and Affiliation - bi-dynamic line graph
```{r echo=FALSE message=FALSE, warning=FALSE}

# Broccatelli, Everett, Koskinen, "Temporal Dynamics in Covert Networks,"
#   Methodological Innovations 9 (2016). http://dx.doi.org/10.1177/2059799115622766


# Make program: findOnes:
findOnes<-function(M) {
  nR<-nrow(M)
  nC<-ncol(M)
  samerows<-matrix(rep((1:nC),nR),nrow=nR,byrow=T)
  samecols<-matrix(rep((1:nR),nC),nrow=nR)
  rows<-M*samecols
  rows<-as.vector(t(rows))
  rows<-rows[which(rows>0)]
  cols<-M*samerows
  cols<-as.vector(t(cols))
  cols<-cols[which(cols>0)]
  # out<-cbind(rows,cols)
  out<-cbind(rows,cols)
  return(out)
}

nTies <- sum(bin_crime>0)
nTies
BD <- matrix(0, nTies, nTies)
nn <- rep("", nTies)
O <- findOnes(bin_crime)

for(k in 1:nTies){
    thisi <- O[k,1]; thisj <- O[k,2]
    nn[k] <- paste("S",thisi,"C",thisj, sep="")
  
}
rownames(BD) <- colnames(BD) <- nn

# Make a program that finds an index in BD:
findPE <-function(O,S,C) {
  # O is the output of findOnes(A)
  # P and E are index numbers of a person (P) and an event (E) in A;
  # S and C are index number of a school (S) and a crime (C) in the binary crime matrix;
  # output is the index number of that node SiCj in BD.
  out <- which((O[,1] == S) & (O[,2] == C))
  return(out)
}

# In BD, add all person-to-person ties WITHIN events
for (j in 1:ncol(bin_crime)) {
  if ((sum(bin_crime[,j])) > 0) {
    this <-outer(bin_crime[,j], bin_crime[,j])
    OJ <- findOnes(this)
    for (k in 1:nrow(OJ)) {
      ti <- findPE(O, OJ[k,1], j)
      tj <- findPE(O, OJ[k,2], j)
      BD[ti,tj] <- 1
    }}}
 
BDssc <- BD # These are the school-to-school ties WITHIN crimes.


 # Now add all temporal crimes (connecting a crime to the next crime, whatever it is) WITHIN each school:
ncA <-ncol(bin_crime)

for (i in 1:nrow(bin_crime)) {
  for (j in 1:(ncA-1)) {
    if((bin_crime[i,j] ==1) & (sum(bin_crime[i,((j+1):ncA)]) > 0)) {
       this <-which(bin_crime[i,((j+1):ncA)] > 0)
       this <- ((j+1):ncA)[this]
       tj <- min(this)
      ti <- findPE(O, i, j)
      tj <- findPE(O, i, tj)
      # print(c(nn[ti], nn[tj]))
      BD[ti,tj] <- 1
    }
  }
}

BD <- BD * !(diag(nTies))

BDcct <- 0 + (BD & ! BDssc) # These are the time flow of crimes (connecting a crime to the next crime, whatever it is) WITHIN each school.
# The output, BD, is the union of BDssc and BDcct 
# Now, to see what we have done, we will plot affiliation matrix bin_crime, and then two different aspects of matrix BD


# # Use a common set of coordinates for points:
# mycoord <- c(-4.71949280739085,-3.14586072764715,-2.62407303974304,-1.47730721356756,-0.151425435354456,1.97849850513494,-0.964220904966701,1.590332185235,2.25986500121079,1.10509628017101,-6.08970027954602,-7.69627995706739,-10.1523730094066,-8.57532837263553,-10.3022233574309,-9.10038623863515,-6.60348146449194,-6.70960597158397,-4.5533934953504,-2.8070976960774)
# mycoord <-matrix(mycoord, ncol=2)
# gplot(BDssc, label=rownames(BDssc), diag=TRUE, coord=mycoord,
#       main="BDssc: School to School ties WITHIN events")
# gplot(BDcct, label=rownames(BDcct), vertex.col="blue", coord=mycoord,
#       main="BDcct: each crime leading to the next, WITHIN schools")
# gplot(BD, label=rownames(BD), vertex.col="green",coord=mycoord,
#       main="BD == bi-dynamic line-graph == union of BDssc | BDcct")


require(sna)
# Plot matrix bin_crime:
my.colors <- c(rep("red",nrow(bin_crime)), rep("blue", ncol(bin_crime)))
gplot(bin_crime, gmode="twomode", 
      label=c(rownames(bin_crime)[15:19],colnames(bin_crime)[15:19]),
      vertex.cex=1.5,
      usearrows=TRUE,
      # label.col = my.colors
      vertex.cols = my.colors
)
# print(bin_crime)

#create crime to school connections
c <- bin_crime %*% t(bin_crime)
gplot(c, edge.lwd = 4*c,
      label=rownames(c)[15:21],
      usearrows=TRUE,
      main = "Crime to School Connections")
# plot school to crime connections
S <- t(bin_crime) %*% bin_crime
gplot(S, edge.lwd = 4*S,
      vertex.col="blue",
      label=rownames(S)[15:31],
      usearrows=TRUE,
      main = "School to Crime Connections")


```


```{r echo=FALSE message=FALSE, warning=FALSE}
# require(tidyverse)
# 
# # To build this network both crime and school columns are joined together to create one long list of nodes. A weights column is also created in a separate grouping of the crime and school columns to show their connecting measures and are labeled "to" and "from". The network is then created by implementing the nodes as rows and columns, and the edge list as the matrix meat. 
# 
# 
# #create node list from the data
# sources <- crime150 %>%
#   distinct(delito) %>%
#   rename(label = delito)
# 
# 
# destinations <- crime150 %>%
#   distinct(escuela_nivel) %>%
#   rename(label = escuela_nivel)
# 
# #put all wanted nodes into one column with id
# nodes <- full_join(sources, destinations, by = "label") %>% rowid_to_column("id")
# 
# 
# #create edges and edge weight from the nodes
# per_route <- crime150 %>%
#   group_by(delito, escuela_nivel) %>%
#   summarise(weight = n()) %>%
#   ungroup()
# # per_route
# 
# # node_mtx <- as.matrix(per_route)
# 
# #link ids
# edges <- per_route %>%
#   left_join(nodes, by = c("delito" = "label")) %>%
#   rename(from = id)
# edges <- edges %>%
#   left_join(nodes, by = c("escuela_nivel" = "label")) %>% 
#   rename(to = id)
# 
# edges <- select(edges, from, to, weight)
# #create a network object
# crime_network <- network(edges, vertex.attr = nodes, matrix.type = "edgelist", ignore.eval = FALSE)
# # plot(crime_network, vertex.cex =3) #, mode = "circle")
# 
# my.colors <- c(rep("red"), rep("blue"))
# gplot(crime_network, gmode="twomode", 
#       label=c(rownames(delito_espesifico),colnames(nivel)),
#       vertex.cex=2,
#       usearrows=TRUE,
#       label.col = my.colors,
#       main = "Crime to School Network Edgelist: Red = 'from', Blue = 'to'",
# )
# 
# install_github("DougLuke/UserNetR")
# 
# 
# net1 <- network::network(crime_network, matrix.type = "adjacency")
# summary(net1)
# gplot(net1, vertex.col = my.colors, main = "Edgelist: Red = 'from', Blue = 'to'") #, displaylabels = TRUE, label.col = my.colors)
# # ?summary
#  
#  ###################################################################################
#  

```

```{r echo=FALSE, message=FALSE, warning=FALSE}


# The crime to schools network only has a degree centrality where the crime with id 1, or "PORTACION DE ARMA DE FUEGO" translated to carrying a firearm, has the highest connections to schools (n = 51), followed by 2, or "ROBO DE ACCESORIOS DE AUTO" translated to stealing autoparts, (n = 29), 3, or "ROBO DE VEHICULO DE SERVICIO PARTICULAR SIN VIOLENCIA" translated to theft of a private vehicle", (n = 20), and 4, or "LESIONES INTENCIONALES POR ARMA BLANCA" translated to stabbings, (n = 17).
# 
# 
# 
# Zeros for the closeness and betweenness centralities are interesting to speculate in regards to school perimiter. Because these crimes don't seem to
# require(sna)
# 
# summary(crime_network)
# 
# print("Degree Centrality: Crimes")
# d.cent <- degree(crime_network) #[1:126, 1:126])
# d.cent
# table(d.cent)
# names(d.cent) <- crime_network %v% "id"
# sort(d.cent)
# # Draw the graph with node sizes proportional to betweenness centralitiy scores:
# gplot(crime_network, label=names(d.cent), vertex.col = my.colors, usearrows=TRUE, vertex.cex = .05 * d.cent) +
# title(main = "Degree Centrality:Crime Network")
# 
# 
# 
# print("Closeness Centrality: Crimes")
# c.cent <- closeness(crime_network, gmode = "digraph")
# c.cent
# names(c.cent) <- crime_network %v% "vertex.names"
# sort(c.cent)
# # note there are all 0s for closeness centrality
# 
# # Draw the graph with node sizes proportional to closeness centralitiy scores:
# gplot(crime_network, label=names(b.cent), vertex.col = my.colors, usearrows=TRUE, vertex.cex = 5 * c.cent) + #label=rownames(bin_crime)
# title(main = "Closeness Centrality:Crime Network")
# 
# print("Betweeness Centrality: Crimes")
# b.cent <- betweenness(crime_network, gmode = "digraph")
# b.cent
# names(b.cent) <- crime_network %v% "vertex.names"
# sort(b.cent)
# # Draw the graph with node sizes proportional to closeness centralitiy scores:
# gplot(crime_network, label=names(b.cent), vertex.col = my.colors, usearrows=TRUE, vertex.cex = .05 * b.cent) +
# title(main = "Betweenness Centrality:crime Network")
# #node c (99.69) and node a (83.19) had the highest betweenness centralities

```



```{r}
##############################################################################################

# memb <-rep(1, 37) # A list of 21 1's.
# memb[b2] <- 2
# memb
# 
# split1 <- rep(1, 37) # The number 1, repeated 21 times
# split1[b2] <- 2 # So split has a "1" for each node in block 1, and a "2" for block 2.
# # Let's see:
# split1

# require(sna)
# crime.2block <- blockmodel(bin_crime, split1, plabels=rownames(bin_crime))
# # Now let's use the "sna" program "plot.blockmodel" to plot what we found:
# # crime.2block
# # crime.2block <- dist(unique(crime.2block$block.membership)) 
# crime.2block <- 
# 
# plot(crime.2block)
# title(main = "Crime CDMX, 2-block split from CONCOR", adj=0)
# # Notice (by looking at the plot) that managers in block 1 don't seek much advice
# #   from managers in block 2. Here are the densities within and between the blocks:
# crime.2block$block.model
# 
# # Compare this to the overall matrix density:
# (sum(bin_crime)) / (37 * 36)
# split4 <-rep(1,37)
# split4[block1.2] <- 2
# split4[block2.1] <- 3
# split4[block2.2] <- 4

# c <- grey(1:length(unique(crime.4block))/length(unique(crime.4block)))
# crime.4block <- blockmodel(bin_crime, split4)
# crime.4block <- dist(unique(crime.4block)) 
# plot(crime.4block, col = c[crime.4block])
# title("Crime, 4-block split", adj=0)



```



## Nodes and Edge List: Create a Network
```{r echo=FALSE message=FALSE, warning=FALSE}
#https://www.jessesadler.com/post/network-analysis-with-r/

# Here we create nodes and edge lists to build our network
require(tidyverse)

#create node list from the data
sources <- crime150 %>%
  distinct(delito) %>%
  rename(label = delito)


destinations <- crime150 %>%
  distinct(escuela_nivel) %>%
  rename(label = escuela_nivel)

#put all wanted nodes into one column with id
nodes <- full_join(sources, destinations, by = "label") %>% rowid_to_column("id")


#create edges and edge weight from the nodes
per_route <- crime150 %>%
  group_by(delito, escuela_nivel) %>%
  summarise(weight = n()) %>%
  ungroup()
per_route

# node_mtx <- as.matrix(per_route)

#link ids
edges <- per_route %>%
  left_join(nodes, by = c("delito" = "label")) %>%
  rename(from = id)
edges <- edges %>%
  left_join(nodes, by = c("escuela_nivel" = "label")) %>% 
  rename(to = id)

edges <- select(edges, from, to, weight)


#create a network object
crime_network <- network(edges, vertex.attr = nodes, matrix.type = "edgelist", ignore.eval = FALSE)
plot(crime_network, vertex.cex =3) #, mode = "circle")

#######################
# library(devtools)
# require(UserNetR)
# d <- degree(delito, gmode = "graph")

install_github("DougLuke/UserNetR")


summary(crime_network)
network.size(crime_network)
gden(crime_network)
component.dist(crime_network)
gtrans(crime_network, mode = "graph")

net1 <- network::network(crime_network, matrix.type = "adjacency")
# summary(net1)
gplot(net1, vertex.col = 2, displaylabels = FALSE)
net2 <- network(crime_network, matrix.type = 'edgelist')
gplot(net2, vertex.col = 2, displaylabels = FALSE)

# summary(net2)



```


##Random Graphs
```{r echo=FALSE message=FALSE, warning=FALSE}
# class(crimes_tidy) # Display the result ## number of random graphs that can be generated given the number of ties
crimes_gnm <- erdos.renyi.game(n=163, 474, type="gnm")

# I want to print two graphs next to each other in the plot screen so ...
op <- par(mfrow = c(1,2)) # Set the plots to have 1 row of plots and 2 plots in that row. ##par = parameters
plot(crimes_gnm, vertex.color="lightblue", main="First random graph")
plot(erdos.renyi.game(163, 474, type="gnm"), vertex.color="beige", main="Second random graph")
par(op)
graph.density(crimes_gnm)
transitivity(crimes_gnm)

#density check
(163*163)/2
474/13284.5


```

#E-I 
```{r echo=FALSE message=FALSE, warning=FALSE}


E_I.basic <-function(M, split, symmetrize=TRUE, point.biserial=TRUE) {
  # See: Everett & Borgatti, Categorical Attribute-Based Centrality: E-I 
  #   and G-F Centrality, Social Networks 34 (2012): 562-569
  #   http://dx.doi.org/10.1016/j.socnet.2012.06.002
  # M is a matrix, split is a vector of group memberships (groups 1 to G)
  # The default is to fuzzy-union M with its transpose.
  #    This works well, e.g., if M <- sp(data) for Closeness
  # However, use symmetrize=FALSE for eigenvector:
  #  e <- eigen(data, which is symmetric)
  #  newMat <- sweep(data, 2, e$vectors[,1], FUN = "*")
  #  newMat <- newMat / e$values[1]
  #  Then run:
  #    E_I.basic(newMat, split, symmetrize=FALSE)
  # This program imposes that the diagonal of the data matrix is 0.
  if(symmetrize == TRUE) {M <- pmax(M, t(M))} # fuzzy union M with its transpose
  if (min(split) == 0) {split <- split + 1}
  nR <- nrow(M); nC <- ncol(M)
  G <- max(split)
  back <- matrix(0,nR,nC)
  for (k in 1:G) {
    w <- which(split == k)
    back[w,w]<- 1
  }
  not.back <- !back
  nid <- !(diag(nR))
  M <- M * nid # Impose that data matrix has only 0's on diagonal.
  back <- back * nid
  not.back <- not.back * nid
  M.I <- sum(M * back)
  M.E <- sum(M * not.back)
  EI.graph <- (M.E - M.I) / (M.E + M.I)
  rsI <- rowSums(M * back)
  rsE <- rowSums(M * not.back)
  EI.node <- (rsE - rsI) / (rsE + rsI)
  nodelevel <- cbind(rsI, rsE, (rowSums(M)), EI.node)
  colnames(nodelevel) <- c("Internal","External","Total","E-I")
  rownames(nodelevel) <- rownames(M)
  # Work on point-biserial correlation:
  a.PB <- sum(back * M)
  c.PB <- sum(not.back * M)
  # t.PB <- table(split)
  # tot.possible.ties.samegroup <- sum(t.PB * (t.PB - 1))
  # tot.possible.ties.matrix <- nR * (nR - 1)
  b.PB <- sum((max(M) - M) * back)
  d.PB <- sum((max(M) - M) * not.back)
  cor.PB <-NA
  point_biserial <- function(a,b,c,d) {
    out <- ((a * d) - (b * c))
    out <- out / sqrt(((a+c) * (b+d) * (a+b) * (c+d)))
    return(out)
  }
  if(point.biserial == TRUE) {
  cor.PB <- point_biserial(a.PB, b.PB, c.PB, d.PB)}
  out <- list(EI.graph=EI.graph, EI.node=nodelevel, point.biserial.graph=cor.PB)
}

E_I.permtest <- function(M, split,howmany=1000, symmetrize=TRUE) {
  # M is a matrix, split is a vector of group memberships (groups 1 to G)
  # Results are for the overall graph, and there are two types:
  #  E_I refers to the E-I index: {that is: E_I.basic(M, split))$EI.graph }
  #  Phi refers to the point-biserial correlation {that is:E_I.basic(M, split))$point.biserial.graph}
  if(symmetrize==TRUE) {M <- pmax(M , t(M))} # fuzzy union M with its transpose
  if (min(split) == 0) {split <- split + 1}
  nR <- nrow(M); nC <- ncol(M)
  G <- max(split)
  obs.E_I <- (res.obs <- E_I.basic(M,split))$EI.graph
  obs.phi <- res.obs$point.biserial.graph
  listEI <- listPhi <- rep(0,howmany)
  for (h in 1:howmany) {
    random.split <- sample(split, nC, replace=FALSE)
    listEI[h] <- (res.h <- E_I.basic(M, random.split))$EI.graph
    listPhi[h] <- res.h$point.biserial.graph
  }
  sorted.list.EI <- sort(listEI)
  sorted.list.Phi <- sort(listPhi)
  q025 <- round(.025*howmany)
  q975 <- round(.975*howmany)
  confint95.EI <- c(sorted.list.EI[q025], sorted.list.EI[q975])
  confint95.Phi <- c(sorted.list.Phi[q025], sorted.list.Phi[q975])
  out<-list(obsEI=obs.E_I, meanEI=mean(listEI), medianEI=median(listEI), confint95EI=confint95.EI, listEI=listEI, obsPhi=obs.phi, meanPhi=mean(listPhi),medianPhi=median(listPhi), confint95Phi=confint95.Phi, listPhi=listPhi)
  return(out)
}

E_I.basic(crime_mtx, 2)

E_I.permtest(mtx_full, 2)
```

```{r}
# I want to print two graphs next to each other in the plot screen so ...
# op <- par(mfrow = c(1,2)) # Set the plots to have 1 row of plots and 2 plots in that row. ##par = parameters
# plot(crimes_tidy, vertex.color="lightblue", main="First random graph")
# plot(erdos.renyi.game(163, 474, type="gnm"), vertex.color="beige", main="Second random graph")
# par(op)
# graph.density(crimes_tidy)
# transitivity(crimes_tidy)
# 

# Demonstrating that, for a large number of nodes (large n),
#  the network will have approximately a Poisson degree distribution.
# (For a symmetric network, the "degree distribution" is the proportion of
#    nodes that have K connections, for each value of K = 0, 1, 2, ...)
## if the binomial is small, the prob is distributed by the Poisson
# op <- par(mfrow = c(1,2))
# # Below, we give a probability that any pair of nodes exhibit a tie (p = .005);
# #    hence, this is a model of type="gnp" or: G(n, p).
# # g <- erdos.renyi.game(n = 1000, .005, type="gnp") #create random graph
# crimes_gnp <- erdos.renyi.game(n=2161, .005, type="gnp")
# op <- par(mfrow = c(1,1))
# plot(crimes_gnp)
# 
# # In the "plot" statement below, "b" means to show both points AND lines.
# op <- par(mfrow = c(1,2)) # split plot screen panels
# plot(degree.distribution(crimes_gnp), type="b", xlab = "Degree", ylab = "Proportion", main = "Observed Degree Distribution")
# 
# 
# # Compare fit of the Poisson distribution, whose mean is the mean degree 
# #   distribution:
# (lambda <- mean(degree(crimes_gnp)))
# k <- 0:14
# Pois <- (exp(-lambda)) * (lambda^k) / factorial(k)
# plot(k, (Pois / sum(Pois)), type="b", xlab="Degree", ylab="Proportion", col="forestgreen", main = "Poisson approximation based on mean degree")
# par(op)
# 
# ## fully connected means any node can reach another in some length
# # Surprisingly, for random graphs, a fully-connected graph tends to emerge
# #  for even very low values of average degree. If p > (2 * ln(n)) / n, then the
# #  random graph is likely to be fully-connected ==> THIS BECOMES MORE TRUE
# #   AS N BECOMES LARGER.
# # For the example above (where we have 12 nodes), the graph is likely to
# #  be fully connected if the average probability of a connection between any two
# #  nodes is greater than
# (my.prob <- (2 *log(163)) / 163) 
# 
# my.N <- 163 ## number of nodes
# my.p <- .06 ## prob that nodes are connected
# list.components <- rep(0, 2161) # This initializes a list with 1,000 entries of 0.
# for (i in 1:2161) { # #what fraction of the 1000 network nodes are fully connected?
#   c <- components(g <- erdos.renyi.game(n = my.N, p = my.p, type = "gnp")) # number of components
#   ##                                                                        asking what is the distribution
#   list.components[i] <- (max(c$csize)) == my.N
#   # c$size is distribution of sizes of components in the graph. 
#   # We demand that (the graph has 1 component that contains all the nodes.)
# }
# # The proportion of all these graphs that are fully connected is:
# (sum(list.components == 1)) / length(list.components)
# 
# # Suppose we had 120 nodes. Then the graph should tend to be connected for
# (my.prob <- (2 * log(120)) / 120)
# # or for my.prob equal to about .08. Run the previous lines again, but using
# my.N <- 120 # and
# my.p <- .08
# 
# 
# transitivity(crimes_gnm)
```


```{r echo=FALSE message=FALSE, warning=FALSE}



```


